---
title: Missing Data Analysis for Mindfulness Intervention Study
author: Kyle M. Lang
date: 2021-09-01
output:
  html_document:
    df_print: paged
---

<style>
  .wideimage{
      overflow-x: scroll;
      white-space: nowrap;
  }

  .wideimage img{
     max-width: none;
  }
</style>

```{r, include = FALSE}
source("setOptions.R")
```

Setup the environment and read in the data.

```{r}
rm(list = ls(all = TRUE)) # Clean the environment

library(mice)

dataDir <- "../../../data/"
plotDir <- "../../figures/"
fn      <- "datList.rds"

datList <- readRDS(paste0(dataDir, fn))
```

First, we need to get a sense of the scope of the missing dat problem. we'll
look at the combined and intervetion-group data separately.

## Combined Data

```{r}
comData <- datList$com
```

What are the variable-wise proportions of missing data?

```{r, max.height = "200px"}
pm <- colMeans(is.na(comData))
pm

range(pm[pm > 0])

hist(pm)
```

How many missing data patterns exist in the data and what do those patterns 
looks like?

<div class="wideimage">
```{r, fig.width = 40, fig.height = 10}
pats <- md.pattern(comData, rotate.names = TRUE)
```
</div>

```{r}
(nPats <- nrow(pats) - 1) # Number of unique missing data patterns
```

We only have `r nPats` missing data patterns, and nonresponse is mostly due to 
attrition.

- The first wave has no missing values
- Several participants who missed the second wave return for the third wave.

What is the covariance coverage?

```{r}
cc <- md.pairs(comData)$rr / nrow(comData)
range(cc[cc < 1])
hist(cc)
```

```{r, eval = FALSE}
for(i in 2:5)
  print(hist(comData[[i]]))

## Columns to exlude from the imputation:
ex <- c(1 : ncol(comData))[-ind]

## Also exclude constants:
ex <- c(ex, which(lev == 1))

## Which column contains the intervention group flag?
tx <- grep("Interventiegroep", colnames(comData))

rMat <- cor(as.matrix(comData[-c(tx, ex)]), use = "pairwise", method = "spearman")
warnings()

## Why are we getting NA correlations with EDS_10_28?
tmp   <- rMat[ , "EDS_10_28"]
probs <- names(tmp)[is.na(tmp)]

for(x in probs) print(table(comData[c(x, "EDS_10_28")]))

table(comData$EDS_10_28)

### NOTE: EDS_10_28 is nearly constant. After accounting for the missing values,
###       EDS_10_28 is exactly constant relative to several other variables.

## Exclude EDS_10_28:
ex <- c(ex, grep("EDS_10_28", colnames(comData)))

## Select the predictors for the imputation model:
preds <- quickpred(data    = comData,
                   mincor  = 0.1,
                   exclude = ex,
                   include = "Interventiegroep",
                   method  = "spearman")

colSums(preds)

meth        <- rep("norm", ncol(comData))
names(meth) <- colnames(comData)

meth[ex]                 <- ""
meth["Interventiegroep"] <- ""
meth[pm == 0]            <- ""

## Impute missing values:
miceOut <- mice(data   = comData,
                m      = 20,
                method = meth,
                pred   = preds,
                maxit  = 10,
                seed   = 235711,
                ridge  = 1.0e-1)

### NOTE: I think we're getting singular matrices because we're using too many
### predictors. Need to check the number of observations and select the
### predictor accordingly.

pdf(paste(plotDir, "mice_traceplots.pdf"), onefile = TRUE)
plot(miceOut)
dev.off()

rowSums(preds)

?quickpred


colMeans(is.na(intData))
colMeans(is.na(comData))

md.pattern(intData)
md.pattern(comData)

intCov <- md.pairs(intData)$rr / nrow(intData)
range(intCov)
hist(intCov)

comCov <- md.pairs(comData)$rr / nrow(comData)
range(comCov)
hist(comCov)
```