---
title: Prepare Variables for Mindfulness Intervention Study
author: Kyle M. Lang
date: 2021-10-05
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
source("setOptions.R")
library(xtable)
```

First, we setup the environment and read in the data that we processed via the
SPSS recoding script.

```{r}
rm(list = ls(all = TRUE))

library(dplyr)

dataDir <- "../../../data/"
fn      <- "recoded_mindfulness_data-20210623.csv"

dat0 <- read.csv2(paste0(dataDir, fn))

head(dat0, 10)
```

---

Now, we convert all categorical variables to factors.

```{r}
dat0$Interventiegroep <- factor(x = dat0$Interventiegroep,
                                labels = c("control", "treatment")
                                )

dat0$marital_status_12 <- factor(x = dat0$marital_status_12,
                                 levels = 1:5,
                                 labels = c("Getrouwd / samenwonend",
                                            "Relatie, maar niet samenwonend",
                                            "Gescheiden / uit elkaar",
                                            "Alleenstaand",
                                            "Anders")
                                 )

dat0$education_12 <- factor(x = dat0$education_12,
                            levels = 1:6,
                            labels = c("Basisonderwijs",
                                       "VMBO / MAVO",
                                       "HAVO / VWO",
                                       "MBO",
                                       "HBO / WO",
                                       "Anders")
                            )

dat0$nationality_12 <- factor(x = dat0$nationality_12,
                              levels = 1:2,
                              labels = c("Nederlandse", "Anders")
                              )

dat0$prev_miscarr_abort_12 <- factor(x = dat0$prev_miscarr_abort_12,
                                     levels = 1:2,
                                     labels = c("Nee", "Ja")
                                     )

dat0$prev_childbirth_12 <- factor(x = dat0$prev_childbirth_12,
                                  levels = 1:2,
                                  labels = c("Nee", "Ja")
                                  )

dat0$planned_pregnancy_12 <- factor(x = dat0$planned_pregnancy_12,
                                    levels = 1:2,
                                    labels = c("Ja", "Nee")
                                    )

dat0$paid_work_12 <-
    factor(x = dat0$paid_work_12,
           levels = 1:3,
           labels = c("Ja",
                      "Nee",
                      "Ja, maar momenteel werk ik langer dan 3 maanden niet")
           )

dat0$alcohol_12 <-
    factor(x = dat0$alcohol_12,
           levels = 1:4,
           labels = c("Nee",
                      "Zo nu en dan, maar niet meer dan 1 glas per week",
                      "Ja, 2-4 glazen per week",
                      "Ja, meer")
           )

dat0$smoking_12 <- factor(x = dat0$smoking_12,
                          levels = 1:2,
                          labels = c("Nee", "Ja")
                          )
```

Check the categorical variables for sparse cells.

```{r, echo = FALSE, results = "asis"}
facVars <- colnames(dat0)[sapply(dat0, is.factor)]

makeTable <- function(target, group, data) {
  tab <- tapply(data[[target]], data[[group]], table, useNA = "ifany")
  tmp <- rbind.data.frame(tab)
  
  rownames(tmp) <- levels(data[[target]])
  
  knitr::kable(tmp, caption = target)
}

for(v in facVars)
  print(makeTable(v, "Interventiegroep", dat0))
```

Several variables are constant within either the treatment or control groups.
There are also several variables with very sparse categories. 

- We'll need to drop some variables from the analysis

  - Those that are constant in one or both groups
  - At least `alcohol_12` and `smoking_12`
  
- We'll need to collapse some categories

   - Those that contain sparse groups
  
---

Check the structure of our data set, so far.

```{r, max.height = "200px"}
str(dat0, list.len = ncol(dat0))
```

The `EndDate_XX` variables are all character. Let's convert them to dates.

```{r}
dateNames       <- grep("EndDate_\\d", colnames(dat0), value = TRUE)
dat0[dateNames] <- lapply(dat0[dateNames], as.Date, format = "%m/%d/%Y")

```

Gestational age is represented as character. We need to convert it to numeric.

```{r}
dat0$Gestational_age_8wpp <- as.numeric(dat0$Gestational_age_8wpp)
```

What's the difference between `Gestational_age_weeks_8wpp` and 
`Gestational_age_8wpp`?

```{r}
tmp           <- dat0[grep("Gestational", colnames(dat0))]
colnames(tmp) <- c("raw", "week")

tmp
tmp[, 1] - tmp[, 2]
```

Is one flavor of gestational age better observed than the other?

```{r}
with(tmp, sum(is.na(raw) & !is.na(week)))
with(tmp, sum(is.na(week) & !is.na(raw)))
```

`Gestational_age_weeks_8wpp` is a truncated version of `Gestational_age_8wpp`.
The former has less missing data than the latter, though. We can only use one, 
so we're better off going with the "weeks" version.

We can also exclude the participant ID.

```{r}
## Drop extraneous variables:
dat0 <- dat0[
    grep("name|age_weeks",
         colnames(dat0),
         ignore.case = TRUE,
         invert = TRUE)
]

head(dat0)
```

`EDS_10` is a screening variable, so we need to drop it from every wave.

```{r}
dat0 <- dat0[grep("EDS_10_\\d", colnames(dat0), invert = TRUE)]
```

Check for any constant columns.

```{r}
lev    <- sapply(dat0, function(x) length(unique(na.omit(x))))
(const <- colnames(dat0)[lev == 1])
```

What are the values of these constants?

```{r}
na.omit(dat0[const])[1, ]
```

Are any of these constant columns mostly missing?

```{r}
colMeans(is.na(dat0[const]))
```

Remove the constant columns.

```{r}
dat0 <- dat0[setdiff(colnames(dat0), const)]
```

---

We only have mindfulness data for the intervention group, so we need to split
the sample.

```{r}
datList <- list()

## Data for only the intervention group:
datList$int <- dat0[dat0$Interventiegroep == "treatment", ]

## Combined data without mindfulness items:
datList$com <- dat0[grep("^TFMQ|^RRQ|^SCS", colnames(dat0), invert = TRUE)]

saveRDS(datList, paste0(dataDir, "datList.rds"))
```


